---
title: "esm242_mini_gp"
author: "Abigail Kirk"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)
```


This problem involves a fish stock that evolves according to $X_{t+1}=X_t+F(X_t)-Y_t$, where the growth function is given by $F(X_t)=rX_t(1-X_t/K)$.  The profits from the fishery in each period are $pY_t-(c/2)*Y_t^2/X_t$ and the catch is proportional to the stock: $Y_t=\alpha X_t$. The objective is to maximize the present value of net benefits (PVNB) from time 0 to time 19 by selecting the catchability coefficient $\alpha$: 

AK notes: Should we / do we need to write another loop to pull individual benefits for SSF / DWF? Would be curious to see how in this case the benefits are currently spread out. But this might be a more interesting question for when we have alpha in their, along with the weightings. 

Agenda for Nate:

- Gut check biological model 
- review mathematics of profit function
- Review current R code
- discuss weighting + how to incorporate into the model
- Assumptions needed?
 
- price is the true reflection of how the people value the fish, but some people eat their fish. would need to make something up
- keep market price, but then keep it as movable parameter + p of consumption, which will never be estimated
 - to ballpark it when uncertain, can we calibrate it to reflect what we see in the real world (i.e. what we see in the current data) 
- estimating costs: we can ignore cost and say that fishers will fish down to some level every single time
- main reason we want convexity in the model, but we have that in the biological model
- hard to estimate weighting parameter and make decisions off of it. 
- what does the shape of our alpha v. beta graph look like?
- if beta = one, then we need to add cost data for SSF to keep convexity 
- what is the cost relative to how much benefit I get. or use cost from some other fishery, 
- assume some linear cost, which c fits the data the best? use captain data, just do simple algebra to find. 
- here can say that this highlights the importance of NaFAA collecting data. 
- Cost Paramater that we can estimate later. 
- we are choosing one alpha across all time periods 
- can use optum 
DD Notes: Adding in another version that optimizes effort, and including a table function to visualize

* model description:
- effort is a choice matrix between two fishers and 10 periods
- the state is the fish stock. In this case we are using a simple schaefer model for stock that both fishers are fishing from
- the yield is the product of the stock and the effort for each fisher.
  * but does this capture that the catch of one impacts the catch of the other?
- Simple individual profit functions are used, optimization is done on the sum of the two functions. 

### Formulas for Biomass and Benefits 

$$
\mathrm{X_{t+1,D}=X_{t,D}+rX_{t,D}(1-\frac{X_{t,D}}{K_{D}}) \\
\text{Where: } X_{0,D}=BI*k_{D} \\ 
\text{   and   } ~~~~~K_{D}=K*k_{D} \\ ~ \\
\Pi_{D}=\sum_{t=0}^{T} \rho^{t} (p (1-\alpha) Y_{t,D} - c_{D} [\frac{Y_{t,D}}{q_DX_{t,D}}]^{2}) \\
\text{Subject to: } Y_{t,D} \lt X_{t,D}}
$$

$$
\mathrm{X_{t+1,S}=X_{t,S}+rX_{t,S}(1-\frac{X_{t,S}}{K_{S}}) \\
\text{Where: } X_{0,S}=BI*k_{S} \\ 
\text{   and   } ~~~~~K_{S}=K*k_{S} \\ ~ \\
\Pi_{S}=\sum_{t=0}^{T} \rho^{t} (p  Y_{t,S} - c_{S} [\frac{Y_{t,S}}{q_SX_{t,S}}]^2) \\
\text{Subject to: } Y_{t,S} \lt X_{t,S}}
$$

$$
\mathrm{BI=60000000 \\
K = 100000000 \\
r=0.45 \\
k_{D}=0.4 \\
k_{S} = 0.6 \\
p= 1.3 \\
c_{D}=0.63 \\
c_{S}= 0.91 \\
q_{D}=0.01 \\
q_{S}=0.03 \\
\delta = 0.05 \\
\rho = \frac{1}{1+\delta} \\
\alpha_{bau}=0.10 \\
\alpha_{opt}=0.507 \\
\beta = 0.7
}
$$


$$
\mathrm{NPV=\beta\Pi_{S}+(1-\beta) \alpha \Pi_D}

$$

```{r}
local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=160000,"local_opts"=local_opts)

harvest_dwf <- function(choice, int_state, alpha, p, c_d, q, r, Kd, discount, period, proportion) {
  benefits=vector(mode="numeric", length=0)
  state=vector(mode="numeric",length=0)
  
  state[1]=int_state*proportion
  benefits[1]=(p*(1-alpha)*choice[1])-(c_d*((choice[1]/(q*state[1]))^2))
  rho=1/(1+discount)
  
  for(i in 2:period){
    state[i]=state[i-1]+(state[i-1]*r*(1-(state[i-1]/Kd)))-choice[i-1]
    benefits[i]=(p*((1-alpha)*choice[i]))-(c_d*((choice[i]/(q*state[i]))^2))
  }
  
  t=seq(from=0,to=period-1)
  pv=rho^t*benefits
  npv=sum(pv)
  return(-npv)
}

constraint_dwf <- function(choice, int_state, alpha, p, c_d, q, r, Kd, discount, period, proportion) {
  
  state=vector(mode="numeric", length=0)
  const=vector(mode="numeric", length=0)
  
  state[1]=int_state*proportion
  const[1]=choice[1]-state[1]
  
  for(i in 2:period) {
    
    state[i]=state[i-1]+(state[i-1]*r*(1-(state[i-1]/Kd)))-choice[i-1]
    const[i]=choice[i]-state[i]
  }
  return(const)
}

harvest_ssf <- function(choice, int_state, p, c_s, q, r, Ks, discount, period, proportion) {
  
  benefits=vector(mode="numeric", length=0)
  state=vector(mode="numeric",length=0)
  
  state[1]=int_state*proportion
  benefits[1]=p*choice[1] - (c_s*((choice[1]/(q*state[1]))^2))
  rho=1/(1+discount)
  
  for(i in 2:period) {
    state[i]=state[i-1]+(state[i-1]*r*(1-(state[i-1]/Ks)))-choice[i-1]
    benefits[i]=p*choice[i] - (c_s*((choice[i]/(q*state[i]))^2))
  }
  
  t=seq(from=0, to=period-1)
  pv=rho^t*benefits
  npv=sum(pv)
  return(-npv)
}


constraint_ssf <- function(choice, int_state, p, c_s, q, r, Ks, discount, period, proportion) {
  
  state=vector(mode="numeric", length=0)
  const=vector(mode="numeric", length=0)
  
  state[1]=int_state*proportion
  const[1]=choice[1]-state[1]
  
  for(i in 2:period){
    
    state[i]=state[i-1]+(state[i-1]*r*(1-(state[i-1]/Ks)))-choice[i-1]
    const[i]=choice[i]-state[i]
  }
  return(const)
}

benefits_fct <- function(choice, Hd_fct, constraintd_fct, Hs_fct, constraints_fct, discount, period, beta){
  
  benefits=vector(mode="numeric",length=0)
  
  rho=1/(1+discount)
  
  
    benefits=(beta*-(nloptr(x0=rep(500000, period),eval_f=Hs_fct,lb=rep(0,period),opts=options,eval_g_ineq=constraints_fct, int_state=60000000,p=1.3,c_s=0.91,q=0.03,r=0.45,Ks=100000000*0.6, discount=0.05,period=20,proportion=0.6)$objective)) + ((1-beta)*choice*-(nloptr(x0 = rep(500000, period),eval_f = Hd_fct,lb = rep(0, period),opts = options,eval_g_ineq = constraintd_fct,int_state = 60000000,alpha = choice,p = 1.3,c_d = 0.63,q = 0.01,r = 0.45, Kd=100000000*0.4,discount = 0.05,period = 20,proportion=0.4)$objective)) 
  
  npv=sum(benefits)
  
  return(-npv)
}

benefits_out=nloptr(x0=0.2,
                    eval_f = benefits_fct,
                    opts = options,
                    lb = 0,
                    ub = 1,
                    Hd_fct = harvest_dwf,
                    constraintd_fct = constraint_dwf,
                    Hs_fct = harvest_ssf,
                    constraints_fct = constraint_ssf,
                    discount = 0.05,
                    period = 20,
                    beta = 0.7
                    )
benefits_out$solution
-benefits_out$objective
0.7*(-H_s_out$objective) + (1-0.7)*(benefits_out$solution*(-H_d_out$objective))
# I=60000000
# K=100000000
# Is=I*0.6
# Id=I*0.4
# Ks=K*0.6
# Kd=K*0.4
# r=0.45
# overall=I+(0.45*I*(1-(I/K)))
# S1=Is+(0.45*Is*(1-(Is/Ks)))
# D1=Id+(0.45*Id*(1-(Id/Kd)))
# S1+D1 == overall
# S1
# D1
```

## Visualizing the above outcome 

### DWF Optimal Solution Table

```{r}
int_state = 60000000
p = 1.3
c_d = 0.63
q = 0.01
r = 0.45
Kd=100000000*0.4
discount = 0.05
period = 20
proportiond=0.4

H_d_out = nloptr(
  x0 = rep(500000, period),
  eval_f = harvest_dwf,
  lb = rep(0, period),
  opts = options,
  eval_g_ineq = constraint_dwf,
  int_state = 60000000,
  alpha = benefits_out$solution,
  p = 1.3,
  c_d = 0.63,
  q = 0.01,
  r = 0.45,
  Kd = 100000000*0.4,
  discount = 0.05,
  period = 20,
  proportion=0.4
)
H_d_out$solution
-H_d_out$objective

dwf_table <- function(choice, int_state, alpha, p, c_d, q, r, Kd, discount, period, proportion) {
  
  benefits=vector(mode="numeric", length=0)
  # choice=vector(mode="numeric",length=0)
  state=vector(mode="numeric",length=0)
  
  state[1]=int_state*proportion
  benefits[1]=(p*(1-alpha)*choice[1])-(c_d*((choice[1]/(q*state[1]))^2))
  
  rho=1/(1+discount)
  
  for(i in 2:period){
    
    state[i]=state[i-1]+(state[i-1]*r*(1-(state[i-1]/Kd)))-choice[i-1]
    
    benefits[i]=(p*((1-alpha)*choice[i]))-(c_d*((choice[i]/(q*state[i]))^2))
  }
  
  t=seq(from=0,to=period-1)
  
  pv=rho^t*benefits
  
  
  return(tibble("Time"=t, "Biomass"=state, "Yield"=choice, "PV"=pv))
}

dwf_df = dwf_table(choice=H_d_out$solution, int_state=int_state, alpha=benefits_out$solution, p=p, c_d=c_d, q=q, r=r, Kd=Kd, discount=discount,period=period,proportion=proportiond) %>% 
  mutate(Fleet = "DWF")
```

### SSF Optimal Solution Table

```{r}
ssf_table <- function(choice, int_state, p, c_s, q, r, Ks, discount, period, proportion) {
  
  benefits=vector(mode="numeric", length=0)
  state=vector(mode="numeric",length=0)
  
  rho=1/(1+discount)
  
  state[1]=int_state*proportion
  benefits[1]=p*choice[1] - (c_s*((choice[1]/(q*state[1]))^2))
  
  for(i in 2:period) {
    
    state[i]=state[i-1]+(state[i-1]*r*(1-(state[i-1]/Ks)))-choice[i-1]
    benefits[i]=p*choice[i] - (c_s*((choice[i]/(q*state[i]))^2))
  }
  
  t=seq(from=0, to=period-1)
  pv=rho^t*benefits
  
  
  return(tibble("Time"=t, "Biomass"=state, "Yield"=choice, "PV"=pv))
}

H_s_out = nloptr(
  x0 = rep(1000000, period),
  eval_f = harvest_ssf,
  lb = rep(0, period),
  opts = options,
  eval_g_ineq = constraint_ssf,
  int_state = 60000000,
  p = 1.3,
  c_s = 0.91,
  q = 0.03,
  r = 0.45,
  Ks = 100000000*0.6,
  discount = 0.05,
  period = 20,
  proportion = 0.6
)

H_s_out$solution
-H_s_out$objective

int_state = 60000000
p = 1.3
c_s = 0.91
q = 0.03
r = 0.45
Ks = 100000000 * 0.6
discount = 0.05
period = 20
proportions = 0.6

ssf_df = ssf_table(choice=H_s_out$solution, int_state=int_state, p=p, c_s=c_s, q=q, r=r, Ks=Ks, discount=discount, period=period, proportion=proportions) %>% 
  mutate(Fleet = "SSF")

total_table= dwf_df %>% 
  rbind(ssf_df)
```

## Plotting Optimal outcome 

```{r}
yield_plot=ggplot(total_table) +
  geom_line(aes(x=Time, y=Yield, color = Fleet), alpha=0.5) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-6, suffix = "M"),
    name = "Yield (Millions of Kg)") +
  labs(title="Optimal Harvest by Fleet when alpha = 0.5") + 
  scale_color_manual(values = c("gold3", "forestgreen")) +
  theme_bw()
yield_plot

# ggsave(filename = "optimized_yield_plot.jpg", plot = yield_plot, path=here::here("Plots"))


# H_s_out$solution
# H_d_out$solution

biomass_plot=ggplot(total_table) +
  geom_line(aes(x=Time, y=Biomass)) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-6, suffix = "M"),
    name = "Stock Biomass (Millions of Kg)") +
  labs(title="Optimal Stock Biomass when alpha = 0.5") + 
  theme_bw()
biomass_plot

# ggsave(filename = "optimized_biomass_plot.jpg", plot = biomass_plot, path=here::here("Plots"))

benefits_plot=ggplot(total_table) +
  geom_line(aes(x=Time, y=PV, color = Fleet), alpha=0.5) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-6, prefix = "$", suffix = "M"),
    name = "Present Value (Millions of Dollars)") +
  labs(title="Optimal Discounted Benefits by Fleet when alpha = 0.5") + 
  scale_color_manual(values = c("gold3", "forestgreen")) +
  theme_bw()
benefits_plot
# ggsave(filename = "optimized_pv_plot.jpg", plot = benefits_plot, path=here::here("Plots"))
```

## Plotting BAU outcome 

```{r}
######################### DWF ##########################
int_state = 60000000
p = 1.3
c_d = 0.63
q = 0.01
r = 0.45
Kd=100000000*0.5
discount = 0.05
period = 20
proportiond=0.5

H_d_og = nloptr(
  x0 = rep(500000, period),
  eval_f = harvest_dwf,
  lb = rep(0, period),
  opts = options,
  eval_g_ineq = constraint_dwf,
  int_state = 60000000,
  alpha = 0.1,
  p = 1.3,
  c_d = 0.63,
  q = 0.01,
  r = 0.45,
  Kd = 100000000*0.4,
  discount = 0.05,
  period = 20,
  proportion=0.4
)
H_d_og$solution
-H_d_og$objective

dwf_df_og = dwf_table(choice=H_d_og$solution, int_state=int_state, alpha=0.1, p=p, c_d=c_d, q=q, r=r, Kd=Kd, discount=discount,period=period,proportion=proportiond) %>% 
  mutate(Fleet = "DWF")

######################### SSF ##########################
H_s_og = nloptr(
  x0 = rep(1000000, period),
  eval_f = harvest_ssf,
  lb = rep(0, period),
  opts = options,
  eval_g_ineq = constraint_ssf,
  int_state = 60000000,
  p = 1.3,
  c_s = 0.91,
  q = 0.03,
  r = 0.45,
  Ks = 100000000*0.6,
  discount = 0.05,
  period = 20,
  proportion = 0.6
)

H_s_og$solution
-H_s_og$objective

int_state = 60000000
p = 1.3
c_s = 0.91
q = 0.03
r = 0.45
Ks = 100000000*0.6
discount = 0.05
period = 20
proportions = 0.6

ssf_df_og = ssf_table(choice=H_s_og$solution, int_state=int_state, p=p, c_s=c_s, q=q, r=r, Ks=Ks, discount=discount, period=period, proportion=proportions) %>% 
  mutate(Fleet = "SSF")

total_table_og = dwf_df_og %>% 
  rbind(ssf_df_og)

################## NAFAA #######################

0.7*(-H_s_og$objective) + (1-0.7)*(0.1*(-H_d_og$objective))

################### PLOTs ######################
yield_plot_og=ggplot(total_table_og) +
  geom_line(aes(x=Time, y=Yield, color = Fleet), alpha=0.5) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-6, suffix = "M"),
    name = "Yield (Millions of Kg)") +
  labs(title="Harvest by Fleet when alpha = 0.1") + 
  scale_color_manual(values = c("gold3", "forestgreen")) +
  theme_bw()
yield_plot_og
yield_plot

# ggsave(filename = "original_yield_plot.jpg", plot = yield_plot_og, path=here::here("Plots"))


biomass_plot_og=ggplot(total_table_og) +
  geom_line(aes(x=Time, y=Biomass)) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-6, suffix = "M"),
    name = "Stock Biomass (Millions of Kg)") +
  labs(title="Stock Biomass when alpha = 0.1") + 
  theme_bw()
biomass_plot_og
biomass_plot
# ggsave(filename = "original_biomass_plot.jpg", plot = biomass_plot_og, path=here::here("Plots"))

benefits_plot_og=ggplot(total_table_og) +
  geom_line(aes(x=Time, y=PV, color = Fleet), alpha=0.5) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-6, prefix = "$", suffix = "M"),
    name = "Present Value (Millions of Dollars)") +
  labs(title="Discounted Benefits by Fleet when alpha = 0.1") + 
  scale_color_manual(values = c("gold3", "forestgreen")) +
  theme_bw()
benefits_plot_og
benefits_plot
# ggsave(filename = "original_pv_plot.jpg", plot = benefits_plot_og, path=here::here("Plots"))
```

