---
title: "esm242_mini_gp"
author: "Abigail Kirk"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)
?nloptr::mlsl
```


This problem involves a fish stock that evolves according to $X_{t+1}=X_t+F(X_t)-Y_t$, where the growth function is given by $F(X_t)=rX_t(1-X_t/K)$.  The profits from the fishery in each period are $pY_t-(c/2)*Y_t^2/X_t$ and the catch is proportional to the stock: $Y_t=\alpha X_t$. The objective is to maximize the present value of net benefits (PVNB) from time 0 to time 19 by selecting the catchability coefficient $\alpha$: 


```{r}

period=11

fish_fct = function(choice, p, c, r, K, period, eff1, eff2, discount) {
  
  state=vector(mode="numeric", length=0)
  # yield=vector(mode="numeric", length=0)
  benefits=vector(mode="numeric", length=0)
  benefits_ssf=vector(mode="numeric", length=0)
  benefits_dwf=vector(mode="numeric", length=0)
  
  rho=1/(1+discount)
  
  yield_matrix = matrix(choice, nrow=2, ncol=period)
  
  state[1]=int_state
  
  benefits_ssf[1]=p*yield[1,1] - c*eff1
  benefits_dwf[1]=p*yield[2,1] - c*eff2
  benefits[1]=benefits_ssf[1]+benefits_dwf[1]
  
  for(i in 2:period){
    state[i]=state[i-1] + (r*state[i-1]*(1-(state[i-1]/K))) - colSums(yield_matrix[,i-1])
    
    benefits_ssf[i]=p*yield[1,i]-c*eff1
    
    benefits_dwf[i]=p*yield[2,i]-c*eff2
    
    benefits[i]=benefits_ssf[i]+benefits_dwf[i]
  }
  
  t=seq(from=0, to=period-1)
  pv=rho^t*benefits
  npv=sum(pv)
  
  return(-npv)
  
}

constraint <- function(choice, p, c, r, K, period, eff1, eff2, discount) {
  
  state=vector(mode="numeric", length=0)
  const=vector(mode="numeric", length=0)
  
  
  yield_matrix = matrix(choice, nrow=2, ncol=period)
  
  # Initialize state and constraint variable; int_state is the time zero value of the state variable; see below for the definition of the constraint variable 
  state[1]=int_state
  
  const[1]=colSums(yield_matrix[,1])-state[1]
  
  
  for(i in 2:period){ # Loop values over the state and define constraint variable
    
    # Let the state evolve; this example matches the state equation for an exhaustible resource 
    state[i]=state[i-1]+(r*state[i-1]*(1-(state[i-1]/K))) - colSums(yield_matrix[,i-1])
    
    # Define constraint variable such that it is less than or equal to zero: if const[i] is leq 0, then choice[i]-state[i] is leq 0 and choice[i] is leq state[i]
    const[i]=colSums(yield_matrix[,i-1])-state[i] 
  }
  return(const)
}

test <- nloptr(x0=rep(1, period),
                 eval_f = fish_fct,
                 lb=rep(0, period), # lower bound, extraction cannot be lower than 0
                 opts = options,
                 int_state=0.5,
                 r=0.1,
                 k=1,
                 eff1= 4, # Define this 
               eff2 = 5, # Define this too (maybe choice?)
                 p=500,
                 c=0.01,
                 discount=.05,
                 eval_g_ineq = constraint
                 )

```




