---
title: "esm242_mini_gp"
author: "Abigail Kirk"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)
?nloptr::mlsl
```


This problem involves a fish stock that evolves according to $X_{t+1}=X_t+F(X_t)-Y_t$, where the growth function is given by $F(X_t)=rX_t(1-X_t/K)$.  The profits from the fishery in each period are $pY_t-(c/2)*Y_t^2/X_t$ and the catch is proportional to the stock: $Y_t=\alpha X_t$. The objective is to maximize the present value of net benefits (PVNB) from time 0 to time 19 by selecting the catchability coefficient $\alpha$: 


```{r}
fish_fct = function(choice, p, c, r, K, period, discount) {
  
  state=vector(mode="numeric", length=0)
  yield=vector(mode="numeric", length=0)
  benefits=vector(mode="numeric", length=0)
  
  yield_matrix = matrix(choice, nrow=2, ncol=10)
  
  state[1]=int_state
  
  benefits[1]=p*(q*state[1]*choice[1]) - c*(choice[1]^2)
  
  
  
}
```




