---
title: "esm242_mini_gp"
author: "Abigail Kirk"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)
?nloptr::mlsl
```


This problem involves a fish stock that evolves according to $X_{t+1}=X_t+F(X_t)-Y_t$, where the growth function is given by $F(X_t)=rX_t(1-X_t/K)$.  The profits from the fishery in each period are $pY_t-(c/2)*Y_t^2/X_t$ and the catch is proportional to the stock: $Y_t=\alpha X_t$. The objective is to maximize the present value of net benefits (PVNB) from time 0 to time 19 by selecting the catchability coefficient $\alpha$: 


```{r}
fish_fct = function(choice, int_state, p, c, r, K, period, eff1, eff2, discount) {
  
  state=vector(mode="numeric", length=0)
  # yield=vector(mode="numeric", length=0)
  benefits=vector(mode="numeric", length=0)
  benefits_ssf=vector(mode="numeric", length=0)
  benefits_dwf=vector(mode="numeric", length=0)
  
  rho=1/(1+discount)
  
  yield_matrix = matrix(choice, nrow=2, ncol=10)
  
  state[1]=int_state
  
  benefits_ssf[1]=p*yield_matrix[1,1, drop = FALSE] - c*eff1
  benefits_dwf[1]=p*yield_matrix[2,1, drop = FALSE] - c*eff2
  benefits[1]=benefits_ssf[1]+benefits_dwf[1]
  
  for(i in 2:period){
    state[i]=state[i-1] + (r*state[i-1]*(1-(state[i-1]/K))) - colSums(yield_matrix[,i-1, drop = FALSE])
    
    benefits_ssf[i]=p*yield_matrix[1,i, drop = FALSE]-c*eff1
    
    benefits_dwf[i]=p*yield_matrix[2,i, drop = FALSE]-c*eff2
    
    benefits[i]=benefits_ssf[i]+benefits_dwf[i]
  }
  
  t=seq(from=0, to=period-1)
  pv=rho^t*benefits
  npv=sum(pv)
  
  return(-npv)
  
}

constraint <- function(choice, int_state, p, c, r, K, period, eff1, eff2, discount) {
  
  state=vector(mode="numeric", length=0)
  const=vector(mode="numeric", length=0)
  
  
  yield_matrix = matrix(choice, nrow=2, ncol=10)
  
  # Initialize state and constraint variable; int_state is the time zero value of the state variable; see below for the definition of the constraint variable 
  state[1]=int_state
  
  const[1]=colSums(yield_matrix[,1, drop = FALSE])-state[1]
  
  
  for(i in 2:period){ # Loop values over the state and define constraint variable
    
    # Let the state evolve; this example matches the state equation for an exhaustible resource 
    state[i]=state[i-1]+(r*state[i-1]*(1-(state[i-1]/K))) - colSums(yield_matrix[,i-1, drop = FALSE])
    
    # Define constraint variable such that it is less than or equal to zero: if const[i] is leq 0, then choice[i]-state[i] is leq 0 and choice[i] is leq state[i]
    const[i]=colSums(yield_matrix[,i, drop = FALSE])-state[i] 
  }
  return(const)
}

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)

period = 10
fishers = 2
total_choices = period * fishers

# choice, p, c, r, K, period, eff1, eff2, discount
test <- nloptr(x0=rep(1, total_choices),
                 eval_f = fish_fct,
                 lb=rep(0, total_choices), # lower bound, extraction cannot be lower than 0
                 opts = options,
                 int_state=0.5,
                 r=0.1,
                 K=1,
                 eff1= 4, # Define this 
               eff2 = 5, # Define this too (maybe as choice?) because it will determine yield
                 p=500,
                 c=0.01,
                 discount=.05,
               period=10,
                 eval_g_ineq = constraint
                 )
## Note that these values are very small because this is currently done between 0-1
options(scipen=999)
test$solution
test$objective

```

AK notes: Should we / do we need to write another loop to pull individual benefits for SSF / DWF? Would be curious to see how in this case the benefits are currently spread out. But this might be a more interesting question for when we have alpha in their, along with the weightings. 

Agenda for Nate:

- Gut check biological model 
- review mathematics of profit function
- Review current R code
- discuss weighting + how to incorporate into the model
- Assumptions needed?


DD Notes: Adding in another version that optimizes effort, and including a table function to visualize


```{r}
effort_fct = function(choice, int_state, p, c1, c2, r, K, period, q1, q2, discount) {
  
  state=vector(mode="numeric", length=0)
  # yield=vector(mode="numeric", length=0)
  benefits=vector(mode="numeric", length=0)
  benefits_ssf=vector(mode="numeric", length=0)
  benefits_dwf=vector(mode="numeric", length=0)
  yield_ssf=vector(mode="numeric", length=0)
  yield_dwf=vector(mode="numeric", length=0)
  yield_total=vector(mode="numeric", length=0)
  
  rho=1/(1+discount)
  
  eff_matrix = matrix(choice, nrow=2, ncol=10)
  
  state[1]=int_state
  
  yield_ssf[1]=q1*state[1]*eff_matrix[1,1]
  yield_dwf[1]=q2*state[1]*eff_matrix[2,1]
  yield_total[1]=yield_ssf[1]+yield_dwf[1]
  
  benefits_ssf[1]=p*yield_ssf[1] - c1*(eff_matrix[1,1, drop=FALSE]^2)
  benefits_dwf[1]=p*yield_dwf[1] - c2*(eff_matrix[2,1, drop=FALSE]^2)
  benefits[1]=benefits_ssf[1]+benefits_dwf[1]
  
  for(i in 2:period){
    state[i]=state[i-1] + (r*state[i-1]*(1-(state[i-1]/K))) - yield_total[i-1]
    
    yield_ssf[i]=q1*state[i]*eff_matrix[1,i]
    benefits_ssf[i]=p*yield_ssf[i] - c1*(eff_matrix[1,i]^2)
    
    
    yield_dwf[i]=q2*state[i]*eff_matrix[2,i]
    benefits_dwf[i]=p*yield_dwf[i] - c2*(eff_matrix[2,i]^2)
    
    yield_total[i]=yield_ssf[i]+yield_dwf[i]
    benefits[i]=benefits_ssf[i]+benefits_dwf[i]
    
  }
  
  t=seq(from=0, to=period-1)
  pv=rho^t*benefits
  npv=sum(pv)
  
  return(-npv)
  
}

constraint_effort <- function(choice, int_state, p, c1, c2, r, K, period, q1, q2, discount) {
  
  state=vector(mode="numeric", length=0)
  state=vector(mode="numeric", length=0)
  # yield=vector(mode="numeric", length=0)
  yield_ssf=vector(mode="numeric", length=0)
  yield_dwf=vector(mode="numeric", length=0)
  yield_total=vector(mode="numeric", length=0)
  const=vector(mode="numeric", length=0)
  
  eff_matrix = matrix(choice, nrow=2, ncol=10)
  # Initialize state and constraint variable; int_state is the time zero value of the state variable; see below for the definition of the constraint variable 
  state[1]=int_state
  
  yield_ssf[1]=q1*state[1]*eff_matrix[1,1]
  yield_dwf[1]=q2*state[1]*eff_matrix[2,1]
  yield_total[1]=yield_ssf[1]+yield_dwf[1]
  
  const[1]=yield_total[1]-state[1]
  
  
  for(i in 2:period){ # Loop values over the state and define constraint variable
    
    # Let the state evolve; this example matches the state equation for an exhaustible resource 
    state[i]=state[i-1]+(r*state[i-1]*(1-(state[i-1]/K))) - yield_total[i-1]
    
    yield_ssf[i]=q1*state[i]*eff_matrix[1,i]
    yield_dwf[i]=q2*state[i]*eff_matrix[2,i]
    yield_total[i]=yield_ssf[i]+yield_dwf[i]
    # Define constraint variable such that it is less than or equal to zero: if const[i] is leq 0, then choice[i]-state[i] is leq 0 and choice[i] is leq state[i]
    const[i]=yield_total[i]-state[i] 
  }
  return(const)
}

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)
period = 10
fishers = 2
total_choices = period * fishers
# choice, p, c, r, K, period, eff1, eff2, discount
test1 <- nloptr(x0=rep(1, total_choices),
                 eval_f = effort_fct,
                 lb=rep(0, total_choices), # lower bound, extraction cannot be lower than 0
                 opts = options,
                 int_state=0.5,
                 r=0.1,
                 K=1,
                 q1=0.1, # Define this 
               q2 = 0.3, # Define this too (maybe choice?)
                 p=500,
                 c1=0.01,
               c2=0.03,
                 discount=.05,
               period=10,
                 eval_g_ineq = constraint_effort
                 )
options(scipen=999)

test1$solution
-test1$objective
```

Table Function

```{r}

effort_table_fct = function(choice, int_state=0.5, r=0.1, K=1, q1=0.1, q2 = 0.3, p=500,c1=0.01,c2=0.03,discount=.05,period=10) {
  
  state=vector(mode="numeric", length=0)
  # yield=vector(mode="numeric", length=0)
  benefits=vector(mode="numeric", length=0)
  benefits_ssf=vector(mode="numeric", length=0)
  benefits_dwf=vector(mode="numeric", length=0)
  yield_ssf=vector(mode="numeric", length=0)
  yield_dwf=vector(mode="numeric", length=0)
  yield_total=vector(mode="numeric", length=0)
  
  rho=1/(1+discount)
  
  eff_matrix = matrix(choice, nrow=2, ncol=10)
  
  state[1]=int_state
  
  yield_ssf[1]=q1*state[1]*eff_matrix[1,1]
  yield_dwf[1]=q2*state[1]*eff_matrix[2,1]
  yield_total[1]=yield_ssf[1]+yield_dwf[1]
  
  benefits_ssf[1]=p*yield_ssf[1] - c1*(eff_matrix[1,1, drop=FALSE]^2)
  benefits_dwf[1]=p*yield_dwf[1] - c2*(eff_matrix[2,1, drop=FALSE]^2)
  benefits[1]=benefits_ssf[1]+benefits_dwf[1]
  
  for(i in 2:period){
    state[i]=state[i-1] + (r*state[i-1]*(1-(state[i-1]/K))) - yield_total[i-1]
    
    yield_ssf[i]=q1*state[i]*eff_matrix[1,i]
    benefits_ssf[i]=p*yield_ssf[i] - c1*(eff_matrix[1,i]^2)
    
    
    yield_dwf[i]=q2*state[i]*eff_matrix[2,i]
    benefits_dwf[i]=p*yield_dwf[i] - c2*(eff_matrix[2,i]^2)
    
    yield_total[i]=yield_ssf[i]+yield_dwf[i]
    benefits[i]=benefits_ssf[i]+benefits_dwf[i]
    
  }
  
  t=seq(from=0, to=period-1)
  pv=rho^t*benefits
  npv=sum(pv)
  
  return(tibble(E_ssf=eff_matrix[1,], E_dwf=eff_matrix[2,], Y_ssf=yield_ssf, Y_dwf=yield_dwf))
}
```







